Real-Time Certified Probabilistic Pedestrian Forecasting

### 2017-05-07 paper/pedestrian_forecasting_iros_2017.tex
Preliminary numbers

\rtext{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians. Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods. Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA). However LTA suffers from not planning for other pedestrian’s positions at future times. \cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input. \cite{Wang2009} and \cite{Emonet2011} both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e. classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model. These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data. \cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene. \cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories. \cite{Alahi2016} uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact. \cite{Alahi2016} outperforms LTA, social forcing based on \cite{Yamaguchi2011},  and IGP from \cite{Trautman2015}. Along with its quick run time, this establishes \cite{Alahi2016} as state-of-the-art for socially based models, and so we our algorithm to theirs.}

@@

\section{Implementation and Experimental results} \label{sec:implementation}
  \rtext{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
  In particular, we must learn the vector fields $\{X_1, \dots, X_n\}$, the posteriors $\Pr( x_0 \mid M)$ and the priors $\Pr(M)$ for $M \in \mathcal{M}$ from the data.
  For the purposes of demonstration, we use the Stanford Drone Dataset \cite{Robicquet2016}.
  More generally, we assume that for a fixed scene we have a database of previously observed trajectories $\{ \hat{x}^1, \dots, \hat{x}^N\}$.
  From this data we tune the parameters of the model ($\{X_1, \dots, X_n\}$, $\Pr( x_0 \mid M)$ and $\Pr(M)$)  appropriately.

  \subsection{Learning the Vector Fields}
 We begin by identifying the number of possible vector-fields.
  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
  %In principal, any clustering algorithm could accomplish this.
\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$. This function measures the distance between the endpoints irrespective of their ordering. The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g. from a storefront to the sidewalk at the edge of a scene. On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models. This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering. We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
  This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

@@

	\caption{\rtext{An illustration of the predictions generated by the various algorithms. In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory. The likelihood of detection is depicted using the virdis color palette. Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent. The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}

	\label{fig:bookstore-1-2}
\end{figure}

@@

	\caption{\rtext{An illustration of the predictions generated by the various algorithms. In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory. The likelihood of detection is depicted using the virdis color palette. Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle. The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}

	\label{fig:death-1-2}
\end{figure}

@@

This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
We implement our model as well as our evaluation code in Python 2.6. 
Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
\rtext{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}

@@

The output distributions of the three algorithms were compared using their integrals over the cells of a regular grid over our domain. 
These integrals are used to visualize the distributions in Figures \ref{fig:bookstore-1-2} and \ref{fig:death-1-2}.
%In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. 
%The ground truth bounding box at each time instance were set equal to the average bounding box dimension centered at the 
%were the average bounding box dimensions in the training data set. 
Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves. \rtext{In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction. We used the area under the curve of the ROC curve as our measure of this. The ROC curve plots the rate of false positive detections against the probability of detecting a pedestrian, or in essence the quality of the detection versus the safety. The area under this curve is a standard measure of the quality of a predictor. Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time. In addition, we used the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time in order to provide a geometric measure of how accurate the predictions are. Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.}
Our predictor behaves better than any of the other compared methods at moderately large $t$.
%, where we capture the behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. 
%Their encoding of velocity results in the prediction lagging behind the actual observed motion. 
This is despite providing the algorithm in \cite{Kitani2012} with the specific final location of each agent.
%Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/the_results.eps}
	\caption{\rtext{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate. We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on social forces at moderate-to-large timescales lowered their performance.Though including social forces into the model is out of the scope of this paper, its incorporation could begin as follows.  In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential \cite{Helbing1995} .}}
	\label{fig:auc_vs_time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/MHD_results.eps}
	\caption{\rtext{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution. The method from \cite{Alahi2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times. At long timescales the MHD to the trajectory of most algorithms converges. Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}}
	\label{fig:mhd_vs_time}
\end{figure}

@@

\rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
This is shown in Table~\ref{tab:time}. 
Our algorithm implementation leveraged its embarrassing parallelism using a relatively naive Python implementation, which split the computation of frames among 18 cores. 
%Our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme. 
The algorithm in \cite{Kitani2012} was timed with minimal modification using the source code provided by the authors.

@@

\section{Conclusion} 
\label{sec:conclusion}
This paper presents a real-time approach to probabilistic pedestrian modeling. 
We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in three critical areas: 1) prediction speed is efficient enabling tight loop real-time control;  2) our error bound on the predictions accurately reflects potential paths for the human limiting false negatives and maximizing the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a wide variety of real world applications. 
These three features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces. 
We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop. \rtext{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.  In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes. It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}

>>>>

\marginnote{\#1.1}
\rtext{On the other hand, many authors have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians. Early work by \cite{Helbing1995} and \cite{Xu2012} describe pedestrians interactions using physically motivated methods. Several models such as \cite{Yamaguchi2011}, \cite{Yi2016}, and \cite{Pellegrini2009} derive their motion models from \cite{Helbing1995} who incorporate collision avoidance through an interaction potential which repels pedestrians from each other, leading to forms of what is called Linear Trajectory Avoidance (LTA). However LTA suffers from not planning for other pedestrian’s positions at future times. \cite{Emonet2011}, \cite{Hospedales2009} , and \cite{Wang2009}  all take optical flow as input. \cite{Wang2009} and \cite{Emonet2011} both use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e. classes of motion within the scene), where \cite{Hospedales2009} substitutes a Markov model. These models are not agent based, and the lack of an explicit motion model limits their predictive power in situations where we know more about the data. \cite{Koppula2016} predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances for humans and objects based on observed objectives within the scene. \cite{Tay2008} , \cite{Wang2008}, and \cite{Trautman2015} create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories. \cite{Alahi2016} uses Long Short-Term Memory to learn pedestrian motion models without making implicit (as in the case of CRFs) or explicit (in the case of social forces) assumptions about the manner in which agents will interact. \cite{Alahi2016} outperforms LTA, social forcing based on \cite{Yamaguchi2011},  and IGP from \cite{Trautman2015}. Along with its quick run time, this establishes \cite{Alahi2016} as state-of-the-art for socially based models, and so we our algorithm to theirs.}

@@

\section{Implementation and Experimental results} \label{sec:implementation}
\marginnote{\#4.1}
  \rtext{Given the model established in the previous section, we describe an implementation to showcase one way the model can be applied to observational data.}
  In particular, we must learn the vector fields $\{X_1, \dots, X_n\}$, the posteriors $\Pr( x_0 \mid M)$ and the priors $\Pr(M)$ for $M \in \mathcal{M}$ from the data.
  For the purposes of demonstration, we use the Stanford Drone Dataset \cite{Robicquet2016}.
  More generally, we assume that for a fixed scene we have a database of previously observed trajectories $\{ \hat{x}^1, \dots, \hat{x}^N\}$.
  From this data we tune the parameters of the model ($\{X_1, \dots, X_n\}$, $\Pr( x_0 \mid M)$ and $\Pr(M)$)  appropriately.

  \subsection{Learning the Vector Fields}
 We begin by identifying the number of possible vector-fields.
  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
  %In principal, any clustering algorithm could accomplish this.
  \marginnote{\#4.2}
\rtext{We then cluster in $\mathbb{R}^4$ using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined by $d((x_1, x_2, x_3, x_4),\mathbf{y}) : \mathbb{R}^4 \times \mathbb{R}^4 \to \mathbb{R} = \min \left\{ d_e((x_1, x_2, x_3, x_4), \mathbf{y}), d_e((x_3, x_4, x_1, x_2), \mathbf{y}) \right\}$, for the euclidean distance $d_e$. This function measures the distance between the endpoints irrespective of their ordering. The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g. from a storefront to the sidewalk at the edge of a scene. On our data, other distance measures from \cite{Morris2009} and \cite{Lee2007} didn't identify coherent motion models. This is partly due to the fact that oftentimes pedestrians would take slightly different routes to get to their destination, and the cumulative effect on the distance measures is enough to hinder the clustering.}
  This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

@@

	\caption{
		\rtext{An illustration of the predictions generated by the various algorithms. In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory. The likelihood of detection is depicted using the virdis color palette. Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] suffer from the inability of their motion model to adequately match the speed of the agent. The algorithm from [13] quickly passes beyond the boundaries of the scene, so the plots are ommitted here.}}
	\marginnote{\#4.7}
	\label{fig:bookstore-1-2}
\end{figure}

@@

	\caption{\rtext{An illustration of the predictions generated by the various algorithms. In this figure, the dot is the start point of the test trajectory, the diamond is the position at time t, and the X is the end of the trajectory. The likelihood of detection is depicted using the virdis color palette. Notice that the Random Walk is imprecise, while the predictions generated by the algorithm in [7] are unable to match the speed of the agent and choose the wrong direction to follow the agent around the circle. The algorithm in [13] passes beyond the boundaries of the scene, and so the plots are ommitted here.}}
	\reversemarginpar\marginnote{\#4.8}
	\label{fig:death-1-2}
\end{figure}

@@

This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
We implement our model as well as our evaluation code in Python 2.6. 
Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
\reversemarginpar\marginnote{\#4.3}
\rtext{We did a 2-fold cross validation by using 20\% of the data for testing and the remainder for training within each scene. We learned separate collections of vector fields and model parameters for each fold on all of the four scenes on which we tested.}
\reversemarginpar\marginnote{\#4.4}
\rtext{Our analysis was conducted on the Coupa, Bookstore, Death Circle, and Gates scenes from the dataset from \cite{Robicquet2016}, with a total of 142 trajectories analyzed.}

@@

The output distributions of the three algorithms were compared using their integrals over the cells of a regular grid over our domain. 
These integrals are used to visualize the distributions in Figures \ref{fig:bookstore-1-2} and \ref{fig:death-1-2}.
%In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. 
%The ground truth bounding box at each time instance were set equal to the average bounding box dimension centered at the 
%were the average bounding box dimensions in the training data set. 
Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
\reversemarginpar\marginnote{\#4.5} \rtext{In our analysis we sought a metric that measured the similarity between the predictor and the ground truth as well as the ``safety'' of the prediction. We used the area under the curve of the ROC curve as our measure of this. The ROC curve plots the rate of false positive detections against the probability of detecting a pedestrian, or in essence the quality of the detection versus the safety. The area under this curve is a standard measure of the quality of a predictor. Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time. In addition, we used the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time in order to provide a geometric measure of how accurate the predictions are. Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.}
Our predictor behaves better than any of the other compared methods at moderately large $t$.
%, where we capture the behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. 
%Their encoding of velocity results in the prediction lagging behind the actual observed motion. 
This is despite providing the algorithm in \cite{Kitani2012} with the specific final location of each agent.
%Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/the_results.eps}
	\caption{\rtext{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate. We sampled the S-LSTM \cite{Alahi2016} model 100 times in order to give them the best chance in this analysis, but their confidence combined with the over-reliance on ourdated pedestrian measurements at moderate-to-large timescales lowered their performance.}}
	\normalmarginpar\marginnote{\#4.9}
	\label{fig:auc_vs_time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/MHD_results.eps}
	\caption{\rtext{A comparison of the Modified Hausdorff Distance from the ground truth of the pedestrian to a 1000 point sample from each distribution. The method from \cite{Alahi2016} does very well at short time scales due to its confidence, but we outperform all other methods at intermediate times. At long timescales the MHD to the trajectory of most algorithms converges. Our method increases positional uncertainty with time, and so \cite{Kitani2012} outperforms us because they know the end point and we do not, and \cite{Alahi2016} places too much significance on the initial positions of other pedestrians at large time scales.}}
	\normalmarginpar\marginnote{\#4.10}
	\label{fig:mhd_vs_time}
\end{figure}

@@

\normalmarginpar\marginnote{\#4.6}
\rtext{The run time per frame for each algorithm was generated using the mean run time for 400 frames, averaged across all of the trajectory data used in the quality analysis.}
This is shown in Table~\ref{tab:time}. 
Our algorithm implementation leveraged its embarrassing parallelism using a relatively naive Python implementation, which split the computation of frames among 18 cores. 
%Our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme. 
The algorithm in \cite{Kitani2012} was timed with minimal modification using the source code provided by the authors.

@@

\section{Conclusion} 
\label{sec:conclusion}
This paper presents a real-time approach to probabilistic pedestrian modeling. 
We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in three critical areas: 1) prediction speed is efficient enabling tight loop real-time control;  2) our error bound on the predictions accurately reflects potential paths for the human limiting false negatives and maximizing the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a wide variety of real world applications. 
These three features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces. 
We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop. \reversemarginpar\marginnote{\#5.1}\rtext{Though including social forces into the model is out of the scope of this paper, incorporation could begin as follows.  In the current motion model with a single vector field, the acceleration of the $i$th agent is given by $\ddot{x}_i = s^2 DX(x_i) \cdot X(x_i)$.  Incorporating a social force $F_{i}$ acting on the $i$th agent can be done by instead asserting $\ddot{x}_i = s^2 DX(x) \cdot X(x) + F_i$.  Usually $F_{i} = \sum_{j} \nabla U( x_j - x_i)$ where $U$ is an interaction potential.} \reversemarginpar\marginnote{\#5.2}\rtext{We also hope to do learning transfer with this model using scene segmentation from \cite{Walker2014}, as well as the semantic context descriptors and routing scores from \cite{Ballan2016} to show how vector fields can be transferred to novel scenes. It appears that the low-order parameterization of our model, and unit-length vector field assumption make it particularly amenable to the methods developed in \cite{Ballan2016}.}
